---
title: "Getting started with wwinference"
description: "A quick start example demonstrating the use of wwinference to jointly fit wastewater and hospital admissions data"
author: "Kaitlyn Johnson"
date: "2024-06-27"
output:
  bookdown::html_vignette2:
    fig_caption: yes
    code_folding: show
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Getting started with wwinference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Quick start

In this quick start, we demonstrate using `wwinference` to specify and fit a
minimal model using daily COVID-19 hospital admissions from a "global" population and
viral concentrations in wastewater from a few "local" wastewater treatment plants,
which come from subsets of the larger population.
This is intended to be used as a reference for those
interested in fitting the `wwinference` model to their own data.

# Package

In this quick start, we also use `dplyr` `tidybayes` and `ggplot2` packages.
These are installed as dependencies when `wwinference` is installed.

```{r load-pkgs, warning=FALSE, message=FALSE}
library(wwinference)
library(dplyr)
library(ggplot2)
library(tidybayes)
```

# Data

The model expects two types of data: daily counts of hospital admissions data
from the larger "global" population, and wastewater concentration
data from wastewater treatment plants whose catchment areas are contained within
the larger "global" population. For this quick start, we will use
simulated data, modeled after a hypothetical US state with 4 wastewater
treatmentplants (also referred to as sites) reporting data on viral
concentrations of SARS-COV-2, processed in 3 different labs, covering about 70%
of the state's population. This simulated data contains daily counts of the
total hospital admissions in a hypothetical US state from September 1, 2023 to
November 29, 2023. It contains wastewater concentration data spanning from
September 1, 2023 to December 1, 2023, with varying sampling frequencies. We
will be using this data to produce a forecast of COVID-19 hospital admissions
as of December 6, 2023. These data are provided as part of the package data.

These data are already in a format that can be used for `wwinference`. For the
hospital admissions data, it contains:
- a date (column `date`): the date of the observation, in this case, the date
the hospital admissions occurred
- a count (column `daily_hosp_admits`): the number of hospital admissions
observed on that day
- a population size (column `state_pop`): the population size covered
by the hospital admissions data, in this case, the size of the theoretical state.

Additionally, we provide the `hosp_data_eval` dataset which contains the
simulated hospital admissions 28 days ahead of the forecast date, which can be
used to evaluate the model.

For the wastewater data, it contains:
- a date (column `date`): the date the sample was collected
- a site indicator (column `site`): the unique identifier for the wastewater treatment plant
that the sample was collected from
- a lab indicator (column `lab`): the unique identifier for the lab where the sample was processed
- a concentration (column `genome_copies_mL`): the measured genome copies per mL from
the sample collected, in natural scale
- a limit of detection (column `lod`): the limit of detection of the assay used to
process the sample, in natural scale
- a site population size (column `site_pop`): the population size covered by the
wastewater catchment area of that site



```{r}
hosp_data <- wwinference::hosp_data
hosp_data_eval <- wwinference::hosp_data_eval
ww_data <- wwinference::ww_data

head(ww_data)
head(hosp_data)
```

# Pre-processing
The user will need to provide data that is in a similar format to the package
data, as described above. This represents the bare minimum required data for a
single location and a single forecast date. We will need to do some
pre-processing to add some additional variables that the model will need to be
able apply features such as outlier exclusion and censoring of values below the
limit of detection.
## Parameters
Get the default parameters from the package. Note that some of these are COVID
specific, others are more general to the model. This is indicated in the
.toml file.
```{r}
params <- get_params(
  system.file("extdata", "example_params.toml",
    package = "wwinference"
  )
)
```

## Wastewater data pre-processing

The `preprocess_ww_data` function adds the following variables to the original
dataset. First, it assigns a unique identifier
the unique combinations of labs and sites, since this is the unit we will
use for estimating the observation error in the reported measurements.
Second it adds a column `below_lod` which is an indicator of whether the
reported concentration is above or below the limit of detection (LOD). If the
point is below the LOD, the model will treat this observation as censored.
Third,  it adds a column `flag_as_ww_outlier` that indicates whether the
measurement is identified as an outlier by our algorithm and the default
thresholds. While the default choice will be to exclude the measurements flagged
as outliers, the user can still choose to include these if they'd like later on.
The user must specify the name of the column containing the
concentration measurements (presumed to be in genome copies per mL) and the
name of the column containing the limit of detection for each measurement. The
function assumes that the original data contains the columns `date`, `site`,
and `lab`, and will return a dataframe with the column names needed to
pass to the downstream model fitting functions.
```{r}
ww_data_preprocessed <- wwinference::preprocess_ww_data(
  ww_data,
  conc_col_name = "genome_copies_per_ml",
  lod_col_name = "lod"
)
```


## Hospital admissions data pre-processing
The `preprocess_hosp_data`  function standardizes the column names of the
resulting datafame. The user must specify the name of the column containing
the daily hospital admissions counts and the population size that the hospital
admissions are coming from (from in this case, a hypothetical US state). The
function assumes that the original data contains the column `date`, and will
return a dataframe with the column names needed to pass to the downstream model
fitting functions.
```{r}
hosp_data_preprocessed <- wwinference::preprocess_count_data(
  hosp_data,
  count_col_name = "daily_hosp_admits",
  pop_size_col_name = "state_pop"
)
```


We'll make some plots of the data just to make sure it looks like what we'd expect:

```{r,  out.width='100%'}
ggplot(ww_data_preprocessed) +
  geom_point(
    aes(
      x = date, y = genome_copies_per_ml,
      color = as.factor(lab_site_name)
    ),
    show.legend = FALSE
  ) +
  geom_point(
    data = ww_data_preprocessed |> filter(genome_copies_per_ml <= lod),
    aes(x = date, y = genome_copies_per_ml, color = "red"),
    show.legend = FALSE
  ) +
  geom_hline(aes(yintercept = lod), linetype = "dashed") +
  facet_wrap(~lab_site_name, scales = "free") +
  xlab("") +
  ylab("Genome copies/mL") +
  ggtitle("Lab-site level wastewater concentration") +
  theme_bw()

ggplot(hosp_data_preprocessed) +
  # Plot the hospital admissions data that we will evaluate against in white
  geom_point(
    data = hosp_data_eval, aes(
      x = date,
      y = daily_hosp_admits_for_eval
    ),
    shape = 21, color = "black", fill = "white"
  ) +
  # Plot the data we will calibrate to
  geom_point(aes(x = date, y = count)) +
  xlab("") +
  ylab("Daily hospital admissions") +
  ggtitle("State level hospital admissions") +
  theme_bw()
```

## Data exclusion
As an optional additional pre-processing step, the user can decide to exclude
certain data points from being included in the model fit procedure. For example,
we recommend excluding the flagged wastewater concentration outliers. To do so
we will use the `indicate_ww_exclusions()` function, which will add the
flagged outliers to the exclude column where indicated.
```{r}
ww_data_to_fit <- wwinference::indicate_ww_exclusions(
  ww_data_preprocessed,
  outlier_col_name = "flag_as_ww_outlier",
  remove_outliers = TRUE
)
```


# Model specification:
We will need to set some metadata to facilitate model specification. In addition
to assigning the forecast date (done above), this also includes:
- forecast date (the date we are making a forecast)
- number of days to calibrate the model for
- number of days to forecast
- specification of the generation interval, in this case for COVID-19
- specification of the delay from infection to the count data, in this case
 from infection to COVID-19 hospital admission


## Calibration time and forecast time
The calibration time represents the number of days to calibrate the count data
to. This must be less than or equal to the number of rows in `hosp_data`. The
forecast horizon represents the number of days from the forecast date to
generate forecasted hospital admissions for. Typically, the hospital admissions
data will not be complete up until the forecast date, and we will refer to the
time between the last hospital admissions data point and the forecast date as
the nowcast time. The model will "forecast" this period, in addition to the
specified forecast horizon.
```{r}
forecast_date <- "2023-12-06"
calibration_time <- 90
forecast_horizon <- 28
```

## Delay distributions
We will pass in some probabiltiy mass functions (pmfs) that are specific to
COVID, and to the delay from infections to hospital admissions, the count
data we are using to fit th emodel. If using a different pathogen or a
different count dataset, these pmfs need to be replaced. We provide them as
package data here. These are both vectors of simplexes (they must sum to 1).

Additionally, the model requires specifying a delay distribution for the
infection feedback term, which essentially describes the delay at which
high incident infections results in  negative feedback on future infections
(due to susceptibility, behavior changes, policies to reduce transmission,
etc.). We by default set this as the generation interval, but this can be
modified as long as the values sum to 1.
```{r}
generation_interval <- wwinference::generation_interval
inf_to_hosp <- wwinference::inf_to_hosp

# Assign infection feedback equal to the generation interval
infection_feedback_pmf <- generation_interval
```
We will pass these to the `model_spec()` function of the `wwinference()` model,
along with the other specified parameters above.


# Precompiling the model
As `wwinference` uses `cmdstan` to fit its models, it is necessary to first
compile the model. This can be done using the compile_model() function.
```{r}
model <- wwinference::compile_model()
```
# Fitting the model
We're now ready to fit the model using the “No-U-Turn Sampler Markov chain
Monte Carlo” method. This is a type of Hamiltonian Monte Carlo (HMC) algorithm
and is the core fitting method used by `cmdstan`. The user can adjust the MCMC
settings (see the documentation for `get_mcmc_options()`),
however this vignette will use
the default parameter settings which includes running 4 parallel chains with
750 warm up iterations, 500 sampling iterations for each chain, a target average
acceptance probability of 0.95 and a maximum tree depth of 12. The user may wish
to adjust these as they are iterating to reduce model run-time or to achieve
better convergence on a real-world use case.

We also pass our preprocessed datasets (`ww_data_to_fit` and
`hosp_data_preprocessed`), specify our model using `get_model_spec()`,
set the MCMC settings using `get_mcmc_options()`, and pass in our
pre-compiled model(`model`) to `wwinference()` where they are combined and
used to fit the model.


```{r warning=FALSE, message=FALSE}
ww_fit <- wwinference::wwinference(
  ww_data_to_fit,
  hosp_data_preprocessed,
  model_spec = get_model_spec(
    forecast_date = forecast_date,
    calibration_time = calibration_time,
    forecast_horizon = forecast_horizon,
    generation_interval = generation_interval,
    inf_to_count_delay = inf_to_hosp,
    infection_feedback_pmf = infection_feedback_pmf,
    params = params
  ),
  mcmc_options = get_mcmc_options(),
  compiled_model = model
)
```

# The `wwinference_fit` object
The `wwinference()` function returns a `wwinference_fit` object which includes
the underlying `CmdStan` object (`fit`), a list of the two sources of input
data (`input_data`), the list of the arguments passed to stan
`stan_data_args`), the list of the MCMC options (`mcmc_options`) passed to
stan.
```{r}
head(ww_fit)
```
# Summarizing and plotting the model fit
The `draws_df` is
intended to provide an easy to work with tibble of posterior draws of
the estimated, nowcasted, and forecasted expected observed hospital admissions
and wastewater concentrations, as well as the latent variables of interest
including the site-level R(t) estimates and the state-level R(t) estimate.
We can then easily plot relevant model outputs against data. This can be useful
to get a sense of if you're model is fitting the data well, and if the
nowcasted/forecast quantities look reasonable.

We can create the plots using the fitting wrapper functions (figures can
also created directly using the `draws_df`).

To generate the `draws_df`, you can just pass the entire `wwinference_fit`
object to the `get_draws_df()` function (or you can pass individual
tibbles required by the function).

```{r generating-figures,  out.width='100%'}
draws_df <- get_draws_df(ww_fit)

plot_hosp <- get_plot_forecasted_counts(
  draws = draws_df,
  count_data_eval = hosp_data_eval,
  count_data_eval_col_name = "daily_hosp_admits_for_eval",
  forecast_date = forecast_date
)
plot_hosp

plot_ww <- get_plot_ww_conc(draws_df, forecast_date)
plot_ww

plot_state_rt <- get_plot_global_rt(draws_df, forecast_date)
plot_state_rt

plot_subpop_rt <- get_plot_subpop_rt(draws_df, forecast_date)
plot_subpop_rt
```

## Diagnostics
While the `wwinference()` function will print out messaging if any of the
diagnostics flags fail, we recommend running diagnostics as a separate
post-processing step on the `CmdStan` fit object. Start by running function
`get_model_diagnostic_flags()` on the stan fit object. Then, we recommend
looking at the `raw_fit_obj$summary()` which will show the diagnostic
metrics for each parameter in the model and can help identify which parameters
are likely to be driving any convergence issues. We have set default thresholds
on the model diagnostics for production-level runs, we recommend adjusting
as needed. For further information on troubleshooting the model diagnostics,
we recommend the (bayesplot tutorial)[https://mc-stan.org/bayesplot/articles/visual-mcmc-diagnostics.html].
```{r}
convergence_flag_df <- wwinference::get_model_diagnostic_flags(
  stan_fit_obj =
    ww_fit$fit$result
)
parameter_diagnostics <- ww_fit$fit$result$summary()
diagnostic_summary <- ww_fit$fit$result$diagnostic_summary(quiet = TRUE)
head(convergence_flag_df)
```

## Fit to only hospital admissions data
The package also has functionality to fit the model without wastewater data.
This can be useful when doing comparisons of the impact the wastewater data
has on the forecast, or as a part of a pipeline where one might choose to
rely on the admissions only model if their are covergence or known data issues
with the wastewater data.

```{r warning=FALSE, message=FALSE}
fit_hosp_only <- wwinference::wwinference(
  ww_data_to_fit,
  hosp_data_preprocessed,
  model_spec = get_model_spec(
    forecast_date = forecast_date,
    calibration_time = calibration_time,
    forecast_horizon = forecast_horizon,
    generation_interval = generation_interval,
    inf_to_count_delay = inf_to_hosp,
    infection_feedback_pmf = infection_feedback_pmf,
    include_ww = FALSE,
    params = params
  ),
  mcmc_options = get_mcmc_options(),
  compiled_model = model
)
```

```{r,  out.width='100%'}
draws_df_hosp_only <- get_draws_df(fit_hosp_only)
plot_hosp_hosp_only <- get_plot_forecasted_counts(
  draws = draws_df_hosp_only,
  count_data_eval = hosp_data_eval,
  count_data_eval_col_name = "daily_hosp_admits_for_eval",
  forecast_date = forecast_date
)
plot_hosp_hosp_only
```
