---
title: "Getting started with wwinference"
description: "A quick start example demonstrating the use of wwinference to jointly fit wastewater and hospital admissions data"
author: "Kaitlyn Johnson"
date: "2024-06-27"
output:
  bookdown::html_vignette2:
    fig_caption: yes
    code_folding: show
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Getting started with wwinference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Quick start

In this quick start, we demonstrate using `wwinference` to specify and fit a
minimal model using daily COVID-19 hospital admissions from a "global" population and
viral concentrations in wastewater from a few "local" wastewater treatment plants,
which come from subsets of the larger population.
This is intended to be used as a reference for those
interested in fitting the `wwinference` model to their own data.

# Package

In this quick start, we also use `dplyr` `tidybayes` and `ggplot2` packages.
These are installed as dependencies when `wwinference` is installed.

```{r}
library(wwinference)
library(dplyr)
library(ggplot2)
library(tidybayes)
```

# Data

The model expects two types of data: daily counts of hospital admissions data
from the larger "global" population, and wastewater concentration
data from wastewater treatment plants whose catchment areas are contained within
the larger "global" population. For this quick start, we will use
simulated data, modeled after a hypothetical US state with 6 wastewater treatment
plants reporting data on viral concentrations of SARS-COV-2, covering about X %
of the state's population. This simulated data is assigned to cover dates
from September 1, 2023 to December 1, 2023. These data are provided as part
of the package data.

These data are already in a format that can be used for `wwinference`. For the
hospital admissions data, it contains:
- a date (column `date`): the date of the observation, in this case, the date
the hospital admissions occurred
- a count (column `daily_hosp_admits`): the number of hospital admissions observed on that day
- a population size (column `state_pop`): the population size covered
by the hospital admissions data, in this case, the size of the theoretical state.

For the wastewater data, it contains:
- a date (column `date`): the date the sample was collected
- a site indicator (column `site`): the unique identifier for the wastewater treatment plant
that the sample was collected from
- a lab indicator (column `lab`): the unique identifier for the lab where the sample was processed
- a concentration (column `genome_copies_mL`): the measured genome copies per mL from
the sample collected, in natural scale
- a limit of detection (column `lod`): the limit of detection of the assay used to
process the sample, in natural scale
- a site population size (column `site_pop`): the population size covered by the
wastewater catchment area of that site

```{r}
hosp_data <- wwinference::hosp_data
ww_data <- wwinference::ww_data
```

# Pre-processing
We will need to set some metadata to facilitate model specification. These include:
- the date on which we are making the forecast
- number of days to calibrate the model for
- number of days to forecast
- specification of the generation interval for COVID-19
- specification of the delay from infection to hospital admissions for COVID-19
- setting the priors and parameters, some of which are COVID-19 specific
```{r}
# Get the parameters
params <- get_params(
  system.file("extdata", "example_params.toml",
    package = "wwinference"
  )
)

# Pre-process the wastewater dataset
ww_data <- ww_data |>
  dplyr::left_join(
    ww_data |>
      dplyr::distinct(lab, site) |>
      dplyr::mutate(
        lab_site = dplyr::row_number()
      ),
    by = c("lab", "site")
  ) |>
  dplyr::mutate(lab_site_name = glue::glue("Site: {site},  Lab:  {lab}"))

# Pre-process the hospital admissions dataset
```

We'll make some plots of the data just to make sure it looks like what we'd expect:

```{r}
# Add plots
ggplot(ww_data) +
  geom_point(
    aes(
      x = date, y = genome_copies_per_ml,
      color = as.factor(lab_site_name)
    ),
    show.legend = FALSE
  ) +
  geom_point(
    data = ww_data |> filter(genome_copies_per_ml <= lod),
    aes(x = date, y = genome_copies_per_ml, color = "red"),
    show.legend = FALSE
  ) +
  geom_hline(aes(yintercept = lod), linetype = "dashed") +
  facet_wrap(~lab_site_name, scales = "free") +
  xlab("") +
  ylab("Genome copies/mL") +
  ggtitle("Lab-site level wastewater concentration") +
  theme_bw()

ggplot(hosp_data) +
  geom_point(
    data = hosp_data_eval, aes(
      x = date,
      y = daily_hosp_admits_for_eval
    ),
    shape = 21, color = "black", fill = "white"
  ) +
  geom_point(aes(x = date, y = daily_hosp_admits)) +
  xlab("") +
  ylab("Daily hospital admissions") +
  ggtitle("State level hospital admissions") +
  theme_bw()
```
